Usability of Mobile Map Applications

### 2014-09-25 usability-study/usability-study.tex
adding files to .gitignore
minor fixes

\subsection{Satisfaction}
\begin{figure}[ht]
\begin{center}
\includegraphics[keepaspectratio, width=.8\textwidth ]{satisfaction.pdf}
\end{center}
\end{figure}
\par
Satisfaction ratings were on a scale from 1 to 10.
\par
Average rating for Apple Maps: $ \approx $ \textit{6.18}
\par
Average rating for Google Maps: $ \approx $ \textit{7.72}

\begin{multicols}{2}
[
\section{Heuristic Evaluation}
Though usability metrics are extremely important in assessing a system, it is also important to heuristically evaluate systems in order to explore \textit{why} they performed the way they do.
My heuristic evaluations of the mobile map applications will be based on the guidelines documents that correspond to the applications.
\subsection{Google Maps}
]
The icons of the Google Maps mobile application comply with the Google Design Guidelines\cite{Google}: simple, modern, and geometric in execution. However, Google Maps completely fails to use any standard iOS icons\cite{Apple}:
\par
\noindent
\includegraphics[width=.45\textwidth]{ios-bar-icons.png}
\par
\noindent
Though it is not required to use these provided iOS icons, the iOS HIG (Human Interface Guidelines) state that it is ``a good idea to use the built-in icons as much as possible because users already know what they mean."\cite{Apple} Balancing consistency in platform with consistency of brand is a challenge, but there is a possibility that adhering more to the iOS HIG could improve usability of Google Maps even more. There was a higher percentage of errors specifically involving clicking wrong icons in the Google Maps application than there was in the Apple Maps application.

@@

\end{multicols}
\clearpage
\begin{multicols}{2}
[
\subsection{Apple Maps}
]
\par
\noindent
\includegraphics[width=.5\textwidth]{apple-maps.png}
\par
\noindent
One guideline from iOS HIG is that important content or functionality should be `elevated':\cite{Apple}
\includegraphics[width=.35\textwidth]{apple-importance.png}
\par
\noindent
This suggests that the icon on the top left-hand corner, which takes you to the directions screen in this case, is the most important. I would have to agree that getting directions is very important functionality, so it seems as if the app is complying to the guidelines. However, a couple of test subjects tried to click on the icon second from the right on the bottom for directions. It is possible that this guideline of important functionality being `elevated' is not intuitive to all users.
\end{multicols}
\medskip
\medskip
\par
The lack of standard iOS icons in the Google Maps app may or may not explain some of the efficiency issues as well as some errors that occurred with the tasks on the mobile Google Maps app. The guideline of important functionality being towards the upper-left corner of the screen may or may not explain some of the efficiency issues as well as errors that occurred with the tasks on the mobile Apple Maps app.
\clearpage
\section{Conclusion}
The usability metrics of the Google Maps mobile application outranked those of the Apple Maps application in almost every way. Though there are some problems with the UI consistency, it does not seem to bother the users (refer back to satisfaction). Still, using non-standard icons or processes isn't without drawbacks. It's loud. It's authorial voice. It's greater cognitive load for users. Anything inconsistent, first or third party is. On an integrated platform like iOS, integrating with the platform offers a lot of benefits, especially when it comes to user experience.

>>>>

\subsection{Satisfaction}
\begin{figure}[ht]
\begin{center}
\includegraphics[keepaspectratio, width=.8\textwidth ]{satisfaction.pdf}
\end{center}
\end{figure}
\par
Satisfaction ratings were on a scale from 1 to 10.
\par
Average rating for Apple Maps: $ \approx $ \textit{6.18}
\par
Average rating for Google Maps: $ \approx $ \textit{7.72}
\par
The results suggest that users are more satisfied with Google Maps.
\begin{multicols}{2}
[
\section{Heuristic Evaluation}
Though usability metrics are extremely important in assessing a system, it is also important to heuristically evaluate systems in order to explore \textit{why} they performed the way they do.
My heuristic evaluations of the mobile map applications are based on the guidelines documents that correspond to the applications.
\subsection{Google Maps--- Icons}
]
The icons of the Google Maps mobile application comply with the Google Design Guidelines\cite{Google}: simple, modern, and geometric in execution. However, Google Maps completely fails to use the standard iOS icons\cite{Apple} (note that the following standard iOS icons are not really in the screen shot of Google Maps on the right) :
\begin{center}
\includegraphics[width=.35\textwidth]{ios-bar-icons.png}
\end{center}
Though it is not required to use these provided iOS icons, the iOS HIG (Human Interface Guidelines) state that it is ``a good idea to use the built-in icons as much as possible because users already know what they mean."\cite{Apple} Balancing consistency in platform with consistency of brand is a challenge, but there is a possibility that adhering more to the iOS HIG could improve usability of Google Maps even more. In this study, there was a higher percentage of errors specifically involving clicking wrong icons in the Google Maps application than there was in the Apple Maps application. 
\begin{center}

@@

\end{center}
\end{multicols}
In general, the users of Google Maps were fairly efficient. There were, however, a couple of anomalies within the set of users of Google Maps who took significantly longer to complete the first task than the rest of the users. These two anomalies both took so long because they initially clicked on the wrong icon. In a future study, it would be interesting to see whether or not  the replacement of the Google icons with iOS standard icons would prevent these kinds of anomalies.
\begin{multicols}{2}
[
\subsection{Apple Maps--- Placement }
]
\begin{center}
\includegraphics[width=.51\textwidth]{apple-maps.png}
\end{center}
\par
\noindent
One guideline from the iOS Human Interface Guidelines is that important content or functionality should be `elevated':\cite{Apple}
\includegraphics[width=.35\textwidth]{apple-importance.png}
\par
\noindent
This suggests that the icon on the top left-hand corner ---which happens to be the directions screen in this app--- is the most important. I would have to agree that getting directions is very important functionality, so it seems as if the app is complying to the guidelines. However, a couple of test subjects tried to click on the icon second from the right on the bottom for directions. It is possible that this guideline of important functionality being `elevated' may not be intuitive to all users.
\end{multicols}
\medskip
\medskip
\par
The lack of standard iOS icons in the Google Maps app may or may not explain some of the efficiency issues as well as some errors that occurred with the tasks on the mobile Google Maps app. The guideline of important functionality being towards the upper-left corner of the screen may or may not explain some of the efficiency issues as well as errors that occurred with the tasks on the mobile Apple Maps app.
\section{Conclusion}
The usability metrics of the Google Maps mobile application outranked those of the Apple Maps application in almost every way. Though there are some problems with the UI consistency, it does not seem to bother the users (refer back to satisfaction). Still, using non-standard icons or processes isn't without drawbacks. It's loud. It's authorial voice. It's greater cognitive load for users. Anything inconsistent, first or third party is. On an integrated platform like iOS, integrating with the platform offers a lot of benefits, especially when it comes to user experience.
